Learning CNN filters from user-drawn image makers for coconut-tree image classification

### 2020-06-08 main.tex
Updates from Overleaf

\title{Learning filters from user provided image makers for convolutional neural networks}

@@

\begin{abstract}
  Identifying the species of trees in a region is essential to identify the type of land use that is taking place. It is also essential to assess and monitor plantation performance and analyze the impact of natural disasters. The manual identification of these trees is tedious, costly, and error-prone, so the application of automatic methods is necessary. Deep learning is quite effective in applications from different domains; however, it usually needs a considerable amount of annotated data so that the models' appropriate training can be done. As deep as the model gets, it needs more data. We propose a training method for feature extractors of convolutional neural networks that need only a minimal set of images and that the CNN designer has more control over. The method uses markers that the user places in relevant regions of some training images to learn the convolutional layers' filters allowing the user to interfere with the training process directly. As this method does not rely on backpropagation or minimization of some cost function to learn the feature extractor filters, it does not require many training images. The method was applied to the problem of determining whether area image patches contain coconut trees or not. The model trained with this method achieved competitive results with state of the art.
\end{abstract}

\section{Introduction}
Deep learning has proven to be applicable to different types of tasks, from image classification to the creation of synthetic data \cite{goodfellow2016deep}. Given the success of the application of deep learning in various areas of knowledge, this technique has also been applied to several tasks of remote sensing. Examples of this technique exist in various problems such as segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}.

The detection of trees is important to produce knowledge about the land use of a region, from the assessment of crop diversity or the performance of plantations in a given season to the assessment of damages after natural disasters such as drilling and flooding. As these plantations can span very large areas, manual detection can be costly and susceptible to errors.

In \cite{fassnacht2016review}, we can see a review of the problem of classification of tree species of aerial images. \cite{puttemans2018comparing} apply deep learning to the problem of detecting coconut trees in aerial images. Aparta et al. \cite{aparna2018cnn} apply convolutional neural networks to count coconut trees on plantations in India. Vargas-Muñoz et al. \cite{8899005} propose an active learning approach to detect coconut trees that seeks to reduce the effort of annotating images. The \cite{8899005} approach is divided into a stage of producing candidate patches for having a coconut tree and a stage where a CNN determines whether a patch has a coconut tree. We propose a different way to produce this classifier.

@@

Our method achieved better precision than conventionally trained VGG, and the difference between precision and recall is much smaller. The Linear SVM did not prove to be a good classifier like the MLP, but, on average, it still got a little better than the VGG-16. Using our method as initialization and training the entire network with backpropagation did not improve the result; this implies that the filters we created are already good filters given the training set. Our feature extractor training method was able to learn good filters from a minimal set of images. VGG has many parameters to learn, so it is susceptible to overfitting with such a small training set. We do not use any technique to handle small data sets to train VGG because we intended to verify our approach viability, and these techniques could also be used with our method. In Figure \ref{fig:ex-classification} we can see examples of misclassified images. In these images, coconut trees appear in different angles, sizes, and shapes, or the boundaries between the coconut tree and the background are tenuous, making it more difficult to identify them. In turn, images that do not contain coconut trees, contain trees that resemble their shape.

>>>>

\title{Learning CNN filters from user-drawn image makers for coconut-tree image classification}

@@

\begin{abstract}
 Identifying species of trees in aerial images is essential for land-use classification, plantation monitoring, and impact assessment of natural disasters. The manual identification of trees in aerial images is tedious, costly, and error-prone, so automatic classification methods are necessary. Convolutional Neural Network (CNN) models have well succeeded in image classification applications from different domains. However, CNN models usually require intensive manual identification to create large training sets. We present a training method for CNN feature extraction that needs only a minimal set of images selected by the user. The method learns the filters of each convolutional layer from user-drawn markers in image regions that discriminate classes, allowing direct expert interference in the training process. It does not rely on optimization based on backpropagation, which considerably reduces the number of required images. We demonstrate its advantages on the binary classification of coconut-tree images against a popular CNN model.
\end{abstract}

\section{Introduction}
Deep learning has proven to be applicable to different tasks, from image classification to data synthesis \cite{goodfellow2016deep}. In remote sensing, the applications have involved segmentation of terrain images \cite{kemker2018algorithms, kampffmeyer2016semantic, hamaguchi2018effective}, building identification \cite{xu2018building, lu2018detecting, liu2018multilevel}, and deforestation monitoring \cite{bragilevsky2017deep}. In this work, we are interested in 
 identifying species of trees from aerial images. The topic is important for land-use classification, plantation monitoring during a season, and damage assessment of natural disasters, such as drilling and flooding. As these plantations can span very large areas, manual identification can be costly, tedious, and susceptible to errors, so automatic methods are necessary.

Classification of tree species in aerial images has been actively investigated~\cite{fassnacht2016review}. In \cite{puttemans2018comparing}, Puttemans et al. present a solution to detect coconut trees based on deep learning. Aparta et al. \cite{aparna2018cnn} apply convolutional neural networks to count coconut trees on plantations in India. Vargas-Muñoz et al. \cite{8899005} propose an active learning approach to detect coconut trees that seeks to reduce the effort of annotating images. The \cite{8899005} approach is divided into a stage of producing candidate patches for having a coconut tree and a stage where a CNN determines whether a patch has a coconut tree. We propose a different way to produce this classifier.

@@

Our method achieved better precision than conventionally trained VGG, and the difference between precision and recall is much smaller. The Linear SVM did not prove to be a good classifier like the MLP, but, on average, it still got a little better than the VGG-16. Using our method as initialization and training the entire network with backpropagation did not improve the result; this implies that the filters we created are already good filters given the training set. Our feature extractor training method was able to learn good filters from a minimal set of images. VGG has many parameters to learn, so it is susceptible to overfitting with such a small training set. We do not use any technique to handle small data sets to train VGG because we intended to verify our approach viability, and these techniques could also be used with our method. In Figure \ref{fig:ex-classification} we can see examples of misclassified images. In these images, coconut trees appear in different angles, sizes, and shapes, or the boundaries between the coconut tree and the background are tenuous, making it more difficult to identify them. In turn, images that do not contain coconut trees, contain trees that resemble their shape. The network designer could put markers in those images in other to improve the feature extractor.
